---
title: "Consultant Project"
author: "Tara Dehdari"
date: "2023-05-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
library(dplyr)
data <- read_csv("C:/Users/tarad/USD GRAD SCHOOL-C/ADS501 Foundations of Data Science/consultancy project/codes/Employee_monthly_salary.csv")

```

## PreProcessing Data

### Missing values

```{r }
missing_counts <- colSums(is.na(data))
print(missing_counts)
```
### Looking for Outliers

```{r}
ggplot(data, aes(x = "", y = GROSS)) +
  geom_boxplot() +
  labs(title = "Boxplot of Gross Income",
       x = "Gross",
       y = "Dollars")

ggplot(data, aes(x = "", y = Net_Pay)) +
  geom_boxplot() +
  labs(title = "Boxplot of Net Income",
       x = "Net",
       y = "Dollars")


```


### Data Description
```{r}
str(Employee_monthly_salary)

num_records <- nrow(Employee_monthly_salary)
num_fields <- ncol(Employee_monthly_salary)

cat("Number of records:", num_records, "\n")
cat("Number of fields:", num_fields, "\n")

#number of females
num_females <- sum(Employee_monthly_salary$Gender == "F")
#number of males
num_males <- sum(Employee_monthly_salary$Gender == "M")
#average salary of females
avg_F_salary <- Employee_monthly_salary %>%
  filter(Gender == "F") %>%
  summarize(avg_F_salary = mean(GROSS)) 
#average salary of males
avg_M_salary <- Employee_monthly_salary %>%
  filter(Gender == "M") %>%
  summarize(avg_M_salary = mean(GROSS)) 

cat("The number of Females are: ", num_females," and the average gross pay for females is:" , avg_F_salary$avg_F_salary, "\n")

cat("The number of Males are: ", num_males," and the average gross pay for males is:" , avg_M_salary$avg_M_salary, "\n")

#age range
age_range <- range(Employee_monthly_salary$Age)

cat("Age range:", age_range[1], "-", age_range[2], "\n")

#number of each person in each age
age_counts <- table(Employee_monthly_salary$Age)

print(age_counts)
```
## Data Exploration 
### Grouping between departments and than job titles in each department to see if women are in leadership postions?
### Barchart of women and men salary between job types
```{r}
#how many jobs
unique_jobs <- unique(Employee_monthly_salary$Designation)
print(unique_jobs)

#filter by top job titles 
top_job_titles <- Employee_monthly_salary %>% 
  count(Designation) %>%
  top_n(20)

#box plot with filtered job titles
ggplot(filter(Employee_monthly_salary, Designation %in% top_job_titles$Designation),
       aes(x = Designation, y = GROSS, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Job Title", y = "Gross Salary", fill = "Gender") +
  ggtitle("Gender-based Salary Distribution by Top Job Titles") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#filter by top job titles 
top_departments <- Employee_monthly_salary %>% 
  count(Department) %>%
  top_n(20)

#box plot with filtered job titles
ggplot(filter(Employee_monthly_salary, Department %in% top_departments$Department),
       aes(x = Department, y = GROSS, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Departments", y = "Gross Salary", fill = "Gender") +
  ggtitle("Gender-based Salary Distribution by Top Departments") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

Barchart of every Designation in the Department LUNA/OPEN Engineering

```{r}
# Filter dataset by Department
luna_open_data <- Employee_monthly_salary %>%
  filter(Department == "LUNA/OPEN Engineering")

# Create the grouped bar chart
ggplot(luna_open_data, aes(x = Designation, y = GROSS, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Job Title", y = "Gross Salary", fill = "Gender") +
  ggtitle("Salary Distribution by Job Titles (LUNA/OPEN Engineering)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


# Filter dataset by Department and Job Title
directory_engineers <- Employee_monthly_salary %>%
  filter(Department == "LUNA/OPEN Engineering" & Designation == "Director..Engineering")

# Count the number of Director..Engineering
count_directory_engineers <- nrow(directory_engineers)

# Print the count
print(count_directory_engineers)

# Get the gender of the Director..Engineering
gender <- directory_engineers$Gender

# Print the gender
print(gender)
```
Barchart of every Designation in the Department AmaTec-Americas TSE

```{r}
# Filter dataset by Department
AMA <- Employee_monthly_salary %>%
  filter(Department == "AmaTec - Americas TSE")

# Create the grouped bar chart
ggplot(AMA, aes(x = Designation, y = GROSS, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Job Title", y = "Gross Salary", fill = "Gender") +
  ggtitle("Salary Distribution by Job Titles (AmaTec - Americas TSE)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


In each department the Directors were graphed to see wether they are male or female

```{r}
# Filter dataset for job titles containing "Director" and count by department and gender
director_counts <- Employee_monthly_salary %>%
  filter(grepl("Director", Designation, ignore.case = TRUE)) %>%
  group_by(Department, Gender) %>%
  summarise(count = n())

# Create the bar chart
ggplot(director_counts, aes(x = Department, y = count, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Department", y = "Count", fill = "Gender") +
  ggtitle("Gender Distribution of Directors in Each Department") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```
Look at the wizards program since female director 
```{r}
# Filter dataset by Department
wizard_program <- Employee_monthly_salary %>%
  filter(Department == "Wizards Program")

# Create the grouped bar chart
ggplot(wizard_program, aes(x = Designation, y = GROSS, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Job Title", y = "Gross Salary", fill = "Gender") +
  ggtitle("Salary Distribution by Job Titles (Wizards Program)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

